### 2022-02-26 manuscript.tex
title change, refs names fixup

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx} % \usepackage[draft]{graphicx}
\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\begin{document}

\title[]{Doublet splitting of fusion alpha-particle driven ion cyclotron emission}

\author{J. W. S. Cook}

\address{UKAEA-CCFE, Culham Science Centre, Abingdon, OX14 3DB, UK}
\ead{james.cook@ukaea.uk}
\vspace{10pt}
\begin{indented}
\item[]Jan 2022
\end{indented}

\begin{abstract}
Ion cyclotron emission (ICE) originating from confined populations of fast ions
  in toroidal fusion plasmas is an important non-invasive, passive diagnostic
  for current and next generation devices. The ability to model the ICE signals
  accurately is an essential step towards inferring the characteristics of
  confined energetic alpha-particles or fast NBI ions and background plasma. In
  this paper, the linear growth rates of the magnetoacoustic cyclotron
  instability, which is the leading explanation for ICE, are calculated in high
  resolution 2D $(k_\perp, k_\parallel)$ space for parameters corresponding to
  JET Pulse No. 26148. These calculations shed further light on the origin of
  doublet-splitting of the cyclotron harmonics as well as the apparent need to
  invoke nonlinear interactions with 1D3V particle-in-cell studies to account
  for lower harmonics with substantial amplitudes: the inclusion of signals of
  ICE from a well resolved range of propagation angles can also account for
  these effects.
\end{abstract}

%
% Uncomment for keywords
%\vspace{2pc}
%\noindent{\it Keywords}: XXXXXX, YYYYYYYY, ZZZZZZZZZ
%
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
%
% Uncomment if a separate title page is required
%\maketitle
% 
% For two-column output uncomment the next line and choose [10pt] rather than [12pt] in the \documentclass declaration
%\ioptwocol
%

%%begin novalidate

\section{Introduction}

Observations of ion cyclotron emission (ICE) have been reported from magnetically confined plasmas first in TFTR\cite{Cauffman1995,Dendy1995} and JET\cite{Cottrell1988,Cottrell1993,Dendy1995}, followed by other tokamaks PDX\cite{Heidbrink1994}, JT-60-U\cite{Ichimura2008}, DIII-D\cite{Heidbrink2011}, ASDEX\cite{DInca2014}, ASDEX-Upgrade\cite{Ochoukov2018}, EAST\cite{Liu2020}, and KSTAR\cite{Chapman2017,Chapman2018}. Stellarators\cite{Saito2013} and FRCs\cite{Nicks2020} have also exhibited this phenomenon. Energetic ion populations that give rise to ICE have their origins in fusion reactions in DT plasmas\cite{Cottrell2000, Cauffman1995}, ICRH\cite{Cottrell2000}, and neutral beam injection\cite{Ochoukov2018}. ICE arising from fusion $\alpha$-particles is a proposed route to further the understanding of burning plasmas in ITER\cite{Dendy2015,McClements2015}.

The foremost explanation of the mechanism responsible for ICE is the
magnetoacoustic cyclotron instability (MCI)\cite{Belikov1976}, which arises from
the free energy present in energetic ions with inverted distributions in
velocity space. Theoretical understanding of ICE is based in analytical
solutions to the linearised Maxwell-Vlasov system of equations\cite{Stix} where
the energetic ion species was modelled as ring-beam for which $f(v_\parallel,
v_\perp) \propto \exp(-(v_\parallel-u_\parallel)^2/ v_{th}^2)\delta(v_\perp -
u_\bot)$ \cite{Dendy1994, McClements1996}. Here $u_\parallel$, $u_\perp$ and
$v_{th}$ are the beam velocity (the signed drift speed along the magnetic
field), ring speed (the speed perpendicular to the magnetic field) and thermal
velocity, respectively. Subsequent studies augmented this understanding via
nonlinear fully-kinetic\cite{Cook2013,Chapman2017} and
hybrid\cite{Carbajal2014,Reman2019} particle-in-cell (PIC) simulations in one
spatial dimension and three vector components for velocities and electromagnetic
field (i.e. 1D3V), and recently 2D3V hybrid-PIC simulations\cite{Carbajal2021}. These
computational approaches operate under similar simplifications as the analytical
treatments in that the plasma is initially spatially homogeneous and periodic
in the spatial domain.
Linearised fluid approaches that incorporate toroidal geometry and
inhomogeneous plasma and magnetic fields have shed light of the eigenmode
structure of compressional Alfv{\'e}n eigenmodes, which like ICE are also fast
Alfv{\'e}n modes driven by energetic ions, in spherical
tokamaks\cite{Sharapov2014, Gorelenkov2016} and conventional aspect ratio
tokamaks \cite{Kolesnichenko2000, Fulop2000, Heidbrink2006}.

The effectiveness of analytical theory to reproduce the observed cyclotron
harmonic peaks in the fluctuations of the magnetic field was first explained
somewhat by 1D PIC simulations. In these simulations the onset of nonlinearity
appeared to imprint the amplitudes of each of the harmonics into the signal as
they are at the end of the linear phase; each mode grows independently until 
growth is terminated for all modes when the fastest mode enters its nonlinear
phase meaning that both the amount of free energy available for instability has
decreased and that the unstable energy ion distribution function has been
perturbed into a more stable state. Analysis shows that only higher cyclotron
harmonics are destabilised\cite{McClements2015} to the MCI, however experimental
observations clearly indicate power at theoretically stable lower
harmonics\cite{Cottrell1988}. In a further step towards accurate recreation of
experimental observations, nonlinear phases of hybrid\cite{Carbajal2014} and
fully kinetic\cite{Cook2013,Chapman2017} PIC simulations went on to show that
wave-wave coupling was responsible for raising the amplitude of lower cyclotron
harmonics. Most recently, hybrid 2D simulations have begun to shed light on the
2D3V dynamics of ICE\cite{Carbajal2021}.

To better enable ICE as a diagnostic for confined fast ions, it is necessary to
further develop our understanding with a range of tools. In this work, I present
the growth rates in 2D wavenumber space of the MCI under the conditions of
outer mid-plane of the JET PTE\cite{Cottrell1993} Pulse No. 26148. These growth
rates are obtained from a dispersion relation solver,
PlasmaDispersionRelations.jl, written in the Julia programming
language\cite{Bezanson2017}. These results narrow the gap between analytical
theory and expensive HPC simulations and can be run readily on desktop computers
in a matter of minutes, further enabling research into this topic. Note that
this study focusses on edge-ICE, named for the region of plasma from which the signal
arises, as opposed to core-ICE\cite{Ochoukov2018}.

In section 2, the computational method is described for calculating the linear growth rates, which are presented for energetic ring-beam alpha-particles with non-zero perpendicular thermal spread in section 3. Results are summarised in section 4.

\section{Computational method}

Solutions to the linearised Maxwell-Vlasov system of equations are found in
parallel and perpendicular wavenumber space, $(k_\parallel, k_\bot)$, for a
ring-beam distribution of energetic minority
alpha-particles and Maxwellian electrons and background thermal majority
deuteron ions. Calculations are made for alpha-particles with ring-beams of two
values of pitch-angle cosine: $u_\parallel=-0.646 u_0 \approx
-\cos(\arcsin(\sqrt{R_0 \sin(\psi_0)/R_{ICE}})) \simeq -1.1 V_A$, which is
associated with the ion-cyclotron emission observed on JET Pulse No. 26148 for
which $R_0=3.46$ m and
$\psi_0=55.2^\circ$ are the major radius and pitch angle at birth, and
$R_{ICE}=4$ m is the major radius of the emission region, see
Figs. 3 and 14 and associated text in Ref. \cite{Cottrell1993}; and
$u_\parallel=0 u_0$ representing a ring distribution with no parallel drift, see
Fig. 1(a) of Ref \cite{Cook2013}. Here, $V_A$ is the Alfv{\'en} speed and $u_0=\sqrt{u_\parallel^2 + u_\perp^2}$ is
the speed of an alpha-particle with an energy of 3.6 MeV. The complex frequency
solutions, $\omega$, along the fast Alfv{\'e}n branch are found on a grid of $(k_\parallel, k_\bot)$.

The plasma parameters approximate those found in the region indicated in Fig. 3
of Ref. \cite{Cottrell1993}, which are associated with the continuous trace
signal of Fig. 2 in the same reference. The electrons and deuterons have
temperatures of 1 keV and the electrons have a number density
of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$. The magnetic field is $2.07$
$\mathrm{T}$, which corresponds a major radius of $4$ m.
The ring-beam distribution function, Eq. 1, is a separable
function of the beam in $v_\parallel$ given by Eq. 2 and the
ring in $v_\perp$ given by Eq. 3.
This distribution function is parameterised by the drift
velocity parallel to the magnetic field $u_\parallel$, the drift speed 
perpendicular to the magnetic
field $u_{\perp}$, and the thermal velocity in both directions $v_{th}$. The
thermal spread of the alpha particles' distribution function 
$v_{th}=v_{\parallel,th}=v_{\perp,th}=u_0/100$. The number density
ratio of alpha-particles to deuterons is $1.5\times10^{-4}$.

\begin{align}
f_\alpha\left(v_\parallel, v_\perp\right) &=
  f_{\alpha,\parallel}\left(v_\parallel\right)f_{\alpha,\perp}\left(v_\perp\right)\\
f_{\alpha,\parallel}\left(v_\parallel\right) &=
  \frac{\exp\left(-\frac{(v_\parallel -
  u_\parallel)^2}{v_{\parallel,th}^2}\right)}{v_{\parallel,th} \sqrt{2
  \pi}}\\
f_{\alpha,\perp}\left(v_\perp\right) &=
  \frac{1}{2\pi}\frac{\exp\left(-\frac{\left(v_\bot -
  u_\bot\right)^2}{v_{\bot,th}^2}\right)}{\left(\frac{\sqrt \pi}{2} v_{\perp,
  th} u_\bot (1 - \mathrm{erf}(-\frac{u_\bot}{v_{\perp,th}})) + \exp(-u_\bot^2 /
  v_{\perp, th}^2) \frac{v_{\perp,th}^2}{2}\right)}\\
\end{align}

Solutions to the linearised Maxwell-Vlasov system of equations are given by the
dispersion relation, Eq. \ref{eq:dettensor}, which is satisfied for sets of
$(\omega,\vec{k})$ given the distribution functions of all the species present.
The hot dielectric tensor\cite{Stix}, $\vec{\vec{\epsilon}}$, Eq. \ref{eq:dielectrictensor} is calculated for each species.

\begin{align}
    0 &= ||\vec{\vec{\epsilon}} + \left(\vec{k} \otimes \vec{k} - \vec{\vec{I}}(\vec{k}\cdot\vec{k})\right) \frac{c^2}{\omega^2}||,\label{eq:dettensor}\\
    \vec{\vec{\epsilon}} &= \vec{\vec{1}} + \sum_s \frac{\Pi_s^2}{\omega}\sum_{n=-\infty}^{\infty}
    \int_0^{\infty}
    \int_{-\infty}^{\infty}\frac{\vec{\vec{S}}_{s,n} 2\pi v_\perp dv_\perp dv_\parallel}{\omega - k_\parallel v_\parallel - n \Omega_s}
    \label{eq:dielectrictensor}
\end{align}

\noindent where

\begin{align}
\vec{\vec{S}}_{s,n} &=
\begin{bmatrix}
\frac{n^2J_n^2}{z_s^2}v_\perp U_s & \frac{inJ_nJ_n'}{z_s}v_\perp U_s & \frac{n
  J_n^2}{z_s} v_\perp W_{s,n} \\
\frac{-inJ_nJ_n'}{z_s}v_\perp U_s & (J_n')^2 v_\perp U_s & -iJ_n J_n' v_\perp
  W_{s,n} \\
\frac{nJ_n^2}{z_s}v_\parallel U_s & iJ_nJ_n'v_\parallel U_s & n J_n^2
  v_\parallel W_{s,n} \\
\end{bmatrix}\label{eq:stenor},\\
U_s &= \frac{\partial f_s}{\partial v_\perp} + \frac{k_\parallel}{\omega}  \left(v_\perp \frac{\partial f_s}{\partial v_\parallel} - v_\parallel \frac{\partial f_s}{\partial v_\perp}\right)\label{eq:uterm},\\
  W_{s,n} &=\left(1- \frac{n\Omega_s}{\omega}\right)\frac{\partial f_s}{\partial v_\parallel} + \frac{n\Omega v_\parallel}{\omega v_\perp}\frac{\partial f_s}{\partial v_\perp}\label{eq:wterm},
\end{align}

\noindent $J_n$ is the $n$th Bessel function of the first kind taking the
argument $z_s=k_\perp v_\perp / \Omega_s$ and $J_n'=\frac{\partial J_n}{\partial z}$.

The integrals over perpendicular velocity required for calculating the
dielectric tensor for Maxwellian distribution functions are given in terms of
modified Bessel functions of the first kind\cite{Stix} for both electrons and
ldeuterons i.e. when $u_\perp =0$. The same perpendicular integrals over the
alpha-particles' ring distribution is calculated numerically via
Gauss-Kronrod\cite{Kronrod1965} quadrature using the GaussGK.jl\cite{QuadGK}
library. An absolute and relative tolerance of $0$ and $10^{-8}$, respectively,
on the value of the integrals is used.

Since all the parallel distribution functions for all species are Maxwellians,
the parallel integrals can be calculated from the well known plasma dispersion
function\cite{Fried1961}, which is laid out in the appendix.

The summation over Bessel function indices in Eq. \ref{eq:dielectrictensor}
continues until the relative change between the current and
previous iteration is less than a tolerance of $10^{-8}$. Root finding of Eq.
\ref{eq:dettensor} in complex frequency plane is performed by an amalgamation of the
Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method,
whereby if the values at the nodes of the Nelder-Mead simplex, which in this
case is just a triangle, have a winding number
of zero and therefore the simplex does not contain a root then the optimisation
procedure follows the Nelder-Mead method, otherwise the simplex bifurcates.
The child simplex that contains the root is used for the next
iteration. The optimisation iteration procedure finishes when the real extent of
the simplex is less than $10^{-4}\, \Omega_i$ and the imaginary extent is less
than $10^{-5}\,\Omega_i$, where $\Omega_i$ is the cyclotron frequency of both
the alpha-particles and deuterons since they share a charge-mass ratio.
This approach is found to be adequately robust and
faster than the Nelder-Mead method or winding number approach used in isolation
of one another. Each 512 by 512 scan in $(k_\perp, k_\parallel)$ takes between 5
and 6 minutes on a laptop.

\section{Linear growth rates}

The growth rates of unstable modes for the case where the pitch-angle cosine
value is zero ($u_\parallel = 0$ in Eq. 1) are plotted in Fig.
\ref{fig:2D_F12_zeropitch}(a) against the associated real frequency (only modes
with growth rates in excess of the solver tolerance are shown i.e. $\Im(\omega)
> 10^{-5} \Omega_i$). In this panel shading indicates parallel wavenumber in
units of $\Omega_i/V_A$. Fig. \ref{fig:2D_F12_zeropitch}(b) plots the same data
in the $(\omega,k_\parallel)$ plane where shading instead indicates growth
rates. The simplicity of the consecutive unstable cyclotron harmonics visible in
panel (a) betrays the complexity of the structure in the $(\omega,k_\parallel)$
plane. Integration over $k_\parallel$ projects the solutions onto the frequency
axis, which loses the structure and produces narrow harmonics. Figure
\ref{fig:2D_imag_zeropitch} gives $(k_\perp, k_\parallel)$ space view of the
same data as Fig. \ref{fig:2D_F12_zeropitch} as well as damped and
normal modes, except shading indicates the growth rates. Overlaid
contours indicate the real frequency in units of ion cyclotron harmonics. Only
narrow regions around the cyclotron harmonics are unstable. Overlaid straight
traces indicate propagation angle. No Doppler shift is present due to the lack
of parallel drift in the alpha-particle ring-beam. The growth rate increases
approximately linearly with cyclotron harmonic number, which is consistent with
the literature\cite{Dendy1994}. Together these figures serve to introduce a 2D
view (growth rates in $(k_\perp, k_\parallel)$ space) of traditionally 1D data
(growth rates against frequency). The importance of a non-zero $u_\parallel$ is
discussed next.

\begin{figure}[ht!]
    % \centering
    \raggedleft
\includegraphics[scale=0.9]{fig1.pdf}
    \caption{(colour online) Unstable fast Alfv{\'e}n wave solutions to the
    linear Maxwell-Vlasov system of equations, Eq. \ref{eq:dettensor}, for
    thermal electrons, deuterons and a ring distribution of energetic
    alpha-particles i.e. when there is zero parallel drift, $u_\parallel=0$ in
    Eq. 1. (a) Upper panel: growth rate as a function of real frequency where shading
    indicates the parallel wavenumber, $k_\parallel$, in units of $\Omega_i/V_A$
    of the complex frequency solution. Consecutive cyclotron harmonics are
    increasingly unstable and are narrow due to the lack of Doppler shift 
    associated with parallel drifts. (b) Lower panel: the same solutions
    as in panel (a) in the $(\omega,k_\parallel)$ plane where shading indicates 
    growth rate. The complex structure of the growth rates that exists in
    wavenumber space is projected into sharp harmonics on the frequency axis.}
    \label{fig:2D_F12_zeropitch}
\end{figure}

\begin{figure}[ht!]
    % \centering
    \raggedleft
\includegraphics[scale=0.9]{fig2.pdf}
    \caption{(colour online) Shading indicates the growth rate of fast
    Alfv{\'e}n waves excited by a ring of fusion born alpha-particles, as a
    function of parallel wavenumber on the ordinate and perpendicular wavenumber
    on the abscissa. No solution is found where shading is absent. The
    alpha-particle distribution function is given by Eq. 1 with
    $u_{\parallel} = 0$. Consecutive straight traces indicate the propagation
    angle with respect to the magnetic field: the horizontal line is exactly
    perpendicular; solid traces fan out at intervals of five degrees; and dashed
    traces indicate intervals of one degree. Curved grey traces indicate the
    location in $(k_\perp, k_\parallel)$ of the cyclotron harmonics (i.e. the
    real component of the complex $\omega$ solutions), which are annotated along
    the left hand side and upper edge of the figure. Frequencies are in units of
    the alpha-particle cyclotron frequency, and wavenumbers are in
    alpha-particle cyclotron frequencies per Alfv{\'e}n speed. Note the up-down
    symmetry due to the lack of parallel drift in the energetic alpha
    particles.}
    \label{fig:2D_imag_zeropitch}
\end{figure}

Fig. \ref{fig:2D_F12_physicalpitch} plots solutions for the physically more
realistic pitch-angle
case ($u_\parallel = -0.646 u_0$) in the same manner as Fig.
\ref{fig:2D_F12_zeropitch} except there is an additional black trace in the
upper panel, which represents the measured JET ICE intensity signal; the continuous
trace from Fig. 2 in Ref. \cite{Cottrell1993}.
The observations of ICE intensity are plotted on a log scale scaled to match
the range of the linear MCI growth rates. This juxtaposition is reasonable since
nonlinear PIC code simulations indicate that all growing modes stop when the
fastest growing mode becomes nonlinear at time $\tau$. Under the assumption
that all modes start growing in unison from identical seed amplitudes then the
$log$ of the amplitude of a mode with growth rate $\gamma$ is proportional to
$\gamma \tau$, allowing for a comparison with observations also on a log scale.
There are several points worth noting about this figure. First, the
growth rate is no longer linear with cyclotron harmonic number. Second,
harmonics are broader due to the non-zero parallel drift that gives rise to
strong Doppler shift with increasing $|k_\parallel|$. Third, doublet splitting
of the $10^{th}$ and $11^{th}$ harmonics is observed: see Fig. 2 of Cottrell 1993 for a single
signal showing multiple doublet splittings; Fig. 3 of Dendy et al 1994 (also
Fig. 9 Dendy et al 1995) shows slices at constant propagation angle that when
all superimposed point towards doublet splitting, albeit under-resolved in
wavenumber space. The following describes how these three features are
connected. Doublet splitting arises due to the dip in growth rate along the 
band of instability where
$k_\parallel$ goes through zero combined with the change in sign of the Doppler
shift of the real frequency as $k_\parallel$ changes sign. The effect is
larger for higher harmonics where the growth rate diminishes to zero for
perpendicular propagation.

This effect can be seen in Fig. \ref{fig:2D_imag_physicalpitch}, which is the
counterpart figure to Fig. \ref{fig:2D_F12_physicalpitch} as Fig.
\ref{fig:2D_imag_zeropitch} is to Fig. \ref{fig:2D_F12_zeropitch},
at e.g. the $12^{th}$ harmonic  near $(k_\parallel \simeq 0\; \Omega_{i}/V_A,
k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $V_A$ is the Alfv{\'e}n speed.
Other possible mechanisms for doublet splitting could be grad-B or curvature
drifts\cite{Cottrell1993,Fulop1998}. Qualitatively, scattered points in Fig.
\ref{fig:2D_F12_physicalpitch}(a) resembles the experimental observations given
by the black trace in that the
signal below the $7^{th}$ cyclotron harmonics is lower and flatter with well
defined peaks and the signal beyond the $7^{th}$ harmonic is a step-change
larger and the harmonics are less easily identifiable. These features cannot be
be recreated with a 1D calculation for these distributions functions. Note that
Fig. \ref{fig:2D_imag_physicalpitch} shows these two different regimes more
clearly. The lower frequency behaviour, where $k_\perp < 7 \Omega_i/V_A$, is
characterised by narrow regions of instability near Doppler shifted cyclotron
harmonics that are stable when $k_\parallel=0$ and unstable at larger
negative parallel wavenumbers. The higher frequency behaviour in the region
$\omega\geq 7\Omega_i$ is dominated by broader regions of instability in
frequency and wavenumber near to Doppler shifted cyclotron harmonics, which have
a complex structure in wavenumber space. Modes numbering 1, 2, 6, 7, 8, 12, and
13 are stable when propagating exactly perpendicular whilst obliquely
propagating modes remain unstable. The magnitude of the Doppler shift of
the growing modes can reach $\approx \Omega_i$ for large propagation angles at 
lower frequencies and also for
oblique angles at higher frequencies; see, for example, the mode
associated with the $15^{th}$ cyclotron harmonic near $k_\parallel \simeq
-\frac{3}{4} \Omega_i/V_A, k_\perp \simeq 14 \Omega_i/V_A$, i.e. a propagation
angle of $93^\circ$. Furthermore, the breadth of the harmonics can be
attributed to the contributions from multiple propagation angles even though the
lower harmonics are largely stable for forward travelling and perpendicular
propagation angles.

\begin{figure}[ht!]
    % \centering
    \raggedleft
\includegraphics[scale=0.9]{fig3.pdf}
    \caption{(colour online) The imaginary component of the frequencies for the
    case with parallel drift $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$.
    Panel (a) plots growth rate against frequency where shading indicates
    parallel wavenumber in units of $\Omega_i/V_A$ and the black trace shows
    the re-scaled ICE signal from JET Pulse \#21648 digitised from Fig. 2 of
    Cottrell et al 1993 (with permission from IOP Publishing), (b) plots the solutions
    in the $(\omega,k_\parallel)$ plane where shading indicates growth rate. Note
    that the Doppler shift associated with the non-zero parallel drift speed
    broadens the unstable peaks in frequency. The $10^{th}$ and $11^{th}$ cyclotron
    harmonics are split into doublets, i.e. they have two peaks.}
    \label{fig:2D_F12_physicalpitch}
\end{figure}

\begin{figure}[ht!]
    % \centering
    \raggedleft
\includegraphics[scale=0.9]{fig4.pdf}
    \caption{(colour online) As in Fig. \ref{fig:2D_imag_zeropitch}, except for
    a ring-beam of alpha particles with a more realistic pitch-angle value
    $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$. The non-zero parallel drift
    speed causes up-down asymmetry; this figure for a pitch-angle cosine of the
    opposite sign, $u_{\parallel} = 0.646 u_0$, is identical except flipped
    about $k_\parallel=0$.}
    \label{fig:2D_imag_physicalpitch}
\end{figure}

\begin{figure}[ht!]
    % \centering
    \raggedleft
\includegraphics[scale=0.9]{fig5.pdf}
    \caption{(colour online) As in Fig. \ref{fig:2D_imag_zeropitch}, except that
    the shading indicates the real part of the smoothed solution normalised by
    the high frequency expression for the fast Alfv{\'e}n wave set out in Eq. 25
    of Ref. \cite{Dendy1994}, which is also Eq. 11 in Ref.
    \cite{McClements1996}. The differences in real frequency for the two sets of
    calculations, a pitch-angle cosine of $-0.646$ and zero, are
    negligible.}
    \label{fig:2D_real_zeropitch}
\end{figure}

\section{Summary}

This paper presents the linear Maxwell-Vlasov dispersion relation of a plasma
consisting of thermal electrons, thermal deuteron fuel ions and a minority
energetic ring-beam of fusion born alpha-particles. These high resolution
$(k_\perp, k_\parallel)$ calculations revisit ICE observations from JET Pulse
No. 26148\cite{Cottrell1993,Dendy1995} and reproduce several important features:
the step change in signal strength from $\omega < 6$ to $\omega > 6$; the
doublet splitting of peaks at $\omega \approx 10\Omega_i$ in Fig.
\ref{fig:2D_F12_physicalpitch}, with origins visible in the fine $(k_\parallel,
k_\perp)$ structure in Fig. \ref{fig:2D_imag_physicalpitch}. The Doppler shift
for large $k_\parallel$ becomes of the order of the alpha-particle cyclotron
frequency, which makes it difficult to determine which parts of the signal come
from which cyclotron harmonic numbers. Further effort is required to account for
the extra doublets visible in the continuous trace associated with JET Pulse No.
26148 in Fig. 2 of Ref. \cite{Cottrell1993}.

These calculations capture the finite perpendicular thermal spread of the alpha
particle distribution function, the integral of which has no known analytically
tractable form. The computational capabilities of the software,
PlasmaDispersionRelations.jl, used to generate the results further enable ICE as
a diagnostic for fusion plasmas since the code is capable of: calculating the
dispersion relation of plasmas consisting of an arbitrary number of species; and
of arbitrary distribution functions, although those with parallel Maxwellians,
like those presented here, are faster. It has been shown that it is possible to
qualitatively recreate signals containing lower harmonics of substantial amplitudes by
invoking 2D linear physics. The presence of ICE signals at lower harmonics in
PIC simulations has been attributed to nonlinear interactions between pairs of
more powerful higher harmonics\cite{Carbajal2014}. This work suggests that
observations of lower ICE harmonics may arise due to 2D
linear physics effects but does not exclude the possiblity that nonlinear 1D
physics\cite{Chapman2018} may also play a part.
There are several important features missing from these calculations. Among them
are the effects of the spatial inhomogeniety of the magnetic field and background
plasma, and the quantisation associatied with toroidal periodicity. These
effects may create an eigenmode structure, not accounted for herein,
associated with the kinetic MCI.
Cottrell et al. \cite{Cottrell1993} does discuss the implications of propagation
angles on ICE signals and indicate that lower harmonics must propagate
increasingly close to exactly perpendicular. These effects remain the subject of
further work to incorporate them in linear and nonlinear studies of ICE.

This work showcases a new high resolution tool for studying ICE and captures the
fine detail of ICE signals, which is challenging to capture with nonlinear PIC
code simulations since they suffer from low resolution in frequency and
wavenumber space due to the phenomenon's transience and computational expense,
respectively. As such, this work represents a complementary tool that builds
upon purely analytical work and nonlinear PIC studies, which in turn enhances
the feasibility of exploiting ICE as a passive diagnostic of confined and
leaving ions in magnetised fusion plasmas.

\section{Acknowledgements}

It is a pleasure to thank R O Dendy and B Chapman-Oplopoiou for useful discussions.

\appendix

\section{}

The plasma dispersion function\cite{Fried1961}

\begin{equation}
\it{Z}_n(z)=\frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}\frac{x^n e^{-x^2}}{x-z}dx,
\end{equation}

\noindent can be calculated via the recurrence relation\cite{Sampoorna2007}

\begin{equation}
{\it Z}_n(z) = z {\it Z}_{n-1}(z) + \begin{cases}
0 & \text{if } n \in \{2k : k \in \mathbb{Z}\}
\\
1 & \text{if } n = 1
\\
\frac{1}{2} & \text{if } n = 3
\\
2^{\frac{1-n}{2}} \Pi_{i=1}^{\frac{n-3}{2}} \left( 2i + 1 \right) & \text{otherwise},
\end{cases}
\end{equation}

\noindent where $n \geq 0$ and the cases $n=1$ and $n=3$ are stated explicitly to avoid confusion. The zeroth term is defined as

\begin{equation}
 {\it Z}_0(z) = i \sqrt{\pi} \mathrm{erfcx} \left(-i z \right)
\end{equation}

\noindent where $\mathrm{erfcx}$ is the scaled complementary complex error function.

\newpage

% \clearpage

%%end novalidate
\begin{thebibliography}{9}

\bibitem{Cauffman1995}
Cauffman S, Majeski R, McClements K G, and Dendy R O 1995 \textit{Nucl. Fusion} \textbf{35} 1597
% , "Alfvenic behaviour of alpha particle driven ion cyclotron emission in TFTR,"

\bibitem{Dendy1995}
Dendy R O {\textit et al} 1995 \textit{Nucl. Fusion} \textbf{35} 1733–1742
%Dendy R O, McClements K G, Lashmore-Davies C, Cottrell G A, Majeski R, and Cauffman S 1995 \textit{Nucl. Fusion} \textbf{35} 1733–1742
%  , “Ion cyclotron emission due to collective instability of fusion prod-ucts and beam ions in TFTR and JET,”

\bibitem{Cottrell1988}
Cottrell G A and Dendy R O 1988 \textit{Phys. Rev. Lett.} \textbf{60} 33-6

\bibitem{Cottrell1993} 
Cottrell G A {\textit et al} 1993 \textit{Nucl. Fusion} \textbf{33}, 31365–1387
%Cottrell G A, Bhatnagar V, Costa O D, Dendy R O, Jacquinot J, McClements K G, McCune D, Nave M, Smeulders P, and Start D 1993 \textit{Nucl. Fusion} \textbf{33}, 31365–1387
% 
% "Ion cyclotron emission measurements during JET deuterium-tritium experiments,"

\bibitem{Heidbrink1994}
Heidbrink W and Sadler G 1994 \textit{Nucl. Fusion} \textbf{34} 535–615
% "The behaviour of fast ions in tokamak experiments,"

\bibitem{Ichimura2008}
Ichimura M {\textit et al} 2008 \textit{Nucl. Fusion} \textbf{48}, 035012
%Ichimura M, Higaki H, Kakimoto S, Yamaguchi Y, Nemoto K, Katano M, Ishikawa M, Moriyama S, and Suzuki T 2008 \textit{Nucl. Fusion} \textbf{48}, 035012
% , "Observation of spontaneously excited waves in the ion cyclotron frequency range on JT-60u,"

\bibitem{Heidbrink2011}
Heidbrink W W et al 2011 \textit{Plasma Phys. Control. Fusion} {\textbf 53} 085028
% "Characterization of off-axis fishbones",

\bibitem{DInca2014}
D’Inca R 2014 \textit{Ion Cyclotron Emission on ASDEX Upgrade PhD Thesis} Ludwig-Maximilians-Universitat

\bibitem{Ochoukov2018}
Ochoukov R et al 2018 \textit{Rev. Sci. Instrum.} 2018 \textbf{89}, 10J101
% , "Observations of core ion cyclotron emission on ASDEX Upgrade tokamak"
% , Bobkov V, Chapman B, Dendy RO and Dunne,M  and Faugel, H  and García-Muñoz,M  and Geiger,B  and Hennequin,P  and McClements,K G  and Moseev,D  and Nielsen,S  and Rasmussen,J  and Schneider,P  and Weiland,M  and Noterdaeme,J-M

\bibitem{Liu2020}
Liu L et al 2020  \textit{Nucl. Fusion} \textbf{60} 044002
% "Ion cyclotron emission driven by deuterium neu- tral beam injection and core fusion reaction ions in EAST,"
% Zhang X, Zhu Y, Qin C, Y Zhao, S Yuan, Y Mao, M Li, K Zhang, J Cheng, L Ai, and Y Cheng

\bibitem{Chapman2017}
Chapman B {\textit et al} 2017 \textit{Nucl. Fusion} \textbf{57} 124004
%Chapman B, Dendy R O, McClements K, Chapman S C, Yun G, Thatipamula S, and Kim M 2017 \textit{Nucl. Fusion} \textbf{57} 124004
% , "Sub-microsecond temporal evolution of edge density during edge localized modes in KSTAR tokamak plasmas inferred from ion cyclotron emission,"

\bibitem{Chapman2018}
Chapman B {\textit et al} 2018 \textit{Nucl. Fusion} \textbf{58} 096027
%Chapman B, Dendy R O, Chapman S C, McClements K G, Yun G S and Thatipamula S G 2018 \textit{Nucl. Fusion} \textbf{58} 096027

\bibitem{Saito2013}
Saito K et al 2013 \textit{Plasma Sci. Technol.} \textbf{15} 209–212
% "Measurement of ion cyclotron emissions by using high-frequency magnetic probes in the LHD," 
%R Kumazawa, T Seki, H Kasahara, G Nomura, F Shimpo, H Igami, M Isobe, K Ogawa, K Toi, M Osakabe, M Nishiura, T Watanabe, S Yamamoto, M Ichimura, and T Mutoh

\bibitem{Nicks2020}
Nicks B S, Magee R, Necas A and Tajima T 2021 \textit{Nucl. Fusion} \textbf{61} 016004
% "Beam-driven ion-cyclotron modes in the scrape-off layer of a field-reversed configuration",

\bibitem{Cottrell2000}
Cottrell G A 2000 \textit{Phys. Rev. Lett.} \textbf{84} 2397-400

\bibitem{Dendy2015}
Dendy R O and McClements K G 2015 \textit{Plasma Physics and Controlled Fusion} \textbf{57} 044002
%  "Ion cyclotron emission from fusion-born ions in large tokamak plasmas: a brief review from JET and TFTR to ITER",

\bibitem{McClements2015}
McClements K G, Dendy R O 2015 \textit{Nucl. Fusion} \textbf{55} 043013
% , "Fast particle-driven ion cyclotron emission (ICE) in tokamak plasmas and the case for an ICE diagnostic in ITER",

\bibitem{Belikov1976}
Belikov V S and Kolesnichenko Ya I 1976 \textit{Sov. Phys. Tech. Phys.} \textbf{20} 1146

\bibitem{Stix}
Stix T H 1992 \textit{Waves in Plasmas. 2nd Edition} (Springer, New York)

\bibitem{Dendy1994}
Dendy R O, Lashmore-Davies C N, McClements K G and Cottrell G A 1994 \textit{Physics of Plasmas} \textbf{1} 1918-28
%  "The excitation of obliquely propagating
% fast Alfvén waves at fusion ion cyclotron
% harmonics"

\bibitem{McClements1996}
McClements K G {\textit et al} 1996 \textit{Phys. Plasmas} \textbf{3} 543
%McClements K G, Dendy R O, Lashmore-Davies C N, Cottrell G A, Cauffman S and Majeski R (1996) \textit{Phys. Plasmas} \textbf{3} 543

\bibitem{Cook2013}
Cook J W S, Dendy R O, Chapman S C 2013 \textit{Plasma Physics and Controlled Fusion} \textbf{55} 065003
% , "Particle-in-cell simulations of the magnetoacoustic cyclotron instability of fusion-born alpha-particles in tokamak plasmas",

\bibitem{Carbajal2014}
Carbajal L, Dendy R O, Chapman S C, and Cook J W S 2014 \textit{Phys. Plasmas} \textbf{21} 012106
% , "Linear and non-linear physics of the magnetoacoustic cyclotron instability of fusion-born ions in relation to ion cyclotron emission,"

\bibitem{Reman2019}
Reman B {\textit et al} 2019 \textit{Nucl. Fusion} \textbf{59} 096013
%Reman B, Dendy R O, Akiyama T, Chapman S C, Cook J W S, Igami  H, Inagaki S, Saito K, and Yun G, 2019 \textit{Nucl. Fusion} \textbf{59} 096013
% “Interpreting observations of ion cyclotron emission from large helical device plasmas with beam-injected ion populations,”

\bibitem{Carbajal2021}
Carbajal L, and Calderón F A 2021 \textit{Phys. Plasmas} \textbf{28} 014505
% , "On the 2D dynamics of the magnetoacoustic cyclotron instability driven by fusion-born ions",

\bibitem{Sharapov2014}
Sharapov S E {\textit et al} 2014 \textit{Phys. Plasmas} \textbf{21} 082501
%Sharapov S E, Lilley M K, Akers R, Ayed N B, Cecconello M, Cook J W S, Cunningham G, and Verwichte E 2014 \textit{Phys. Plasmas} \textbf{21} 082501
% , "Bi-directional Alfvén cyclotron instabilities in the mega-amp spherical tokamak"

\bibitem{Gorelenkov2016}
Gorelenkov N N 2016 \textit{New J. Phys.} \textbf{18} 105010

\bibitem{Kolesnichenko2000}
Kolesnichenkoo Ya I, F{\"u}l{\"o}p T, Lisak M and Anderson D 1998 \textit{Nucl. Fusion} {\textbf 38} 1871
% , "Localized fast magnetoacoustic eigenmodes in tokamak plasmas"

\bibitem{Fulop2000}
F{\"u}l{\"o}p T, Lisak M, Kolesnichenko Ya I, and Anderson D, 2000 \textit{Phys. Plasmas} \textbf{7} 1479–1486
% “The radial and poloidal localization of fast magnetoacoustic eigenmodes in tokamaks,”

\bibitem{Heidbrink2006}
Heidbrink W W {\textit et al} A 2006 \textit{Nucl. Fusion} \textbf{46} 324
%Heidbrink W W, Fredrickson E, Gorelenkov N N, Rhodes T, Zeeland M A 2006 \textit{Nucl. Fusion} \textbf{46} 324
% Observation of compressional Alfvén eigenmodes (CAE) in a conventional tokamak.

\bibitem{Bezanson2017}
Bezanson J, Edelman A, Karpinski S, and Shah V B 2017 \textit{SIAM review} \textbf{59} 1
% Julia: A fresh approach to numerical computing,

\bibitem{Fredrickson2019}
Fredrickson E D {\textit et al} 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
%Fredrickson E D, Gorelenkov  N N, Bell R E, Diallo A, LeBlanc B P, Podestà  M, and NSTX Team 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
% "Emission in the ion cyclotron range of frequencies (ICE) on NSTX and NSTX-U",

\bibitem{Kronrod1965}
Kronrod A S 1965 \textit{Nodes and weights of quadrature formulas. Sixteen-place tables} (New York: Consultants Bureau (Authorized translation from the Russian))

\bibitem{QuadGK}
Johnson S G \textit{QuadGK.jl: Gauss-Kronrod integration in Julia} https://github.com/JuliaMath/QuadGK.jl

\bibitem{Nelder1965}
Nelder J A and Mead R 1965 \textit{Computer Journal} \textbf{7} 4
% , A simplex method for function minimization,

\bibitem{Fulop1998}
F{\"u}l{\"o}p T and Lisak M 1998 \textit{Nucl. Fusion} \textbf{38} 761

\bibitem{Fried1961}
Fried B D and Conte S C 1961 \textit{The Plasma Dispersion Function} (Academic Press, New York)

\bibitem{Sampoorna2007}
Sampoorna M, Nagendraa K N, Frisch H 2007 \textit{Journal of Quantitative Spectroscopy \& Radiative Transfer} \textbf{104} 71-85

\end{thebibliography}

\end{document}
% Splitting of spectral peaks
% ‘G. A. Cottrell, V. P. Bhatnagar, 0. da Costa, R. 0. Dendy, A. Edwards, J. Jacquinot, F. Nave, M. Schmid, A. Sibley, P. Smeulders, and D. F. H. Start, in Proceedings of the 19th European Conference on Controlled Fusion and Plasma Heating, Innsbruck, Austria, 1992, edited by W. Freysinger, K. Lackner, R. Schrittwieser, and W. Lindinger (European Physical Soci- ety, Petit-Lancy, Switzerland, 1990). Part I, Vol. 16C. p. 327.
% ‘The JET Team, Nucl. Fusion 32, 187 (1992).
% 7C. A. Cottrell, V. P. Bhatnagar, 0. da Costa, R. 0. Dendy, J. Jacquinot, K.
% G. McClements, D. C. McCune, M. E F. Nave, P. Smeulders, and D. F. H.
% Start, Nucl. Fusion 33, 1365 (1993).